---
title: "Exercise Solution: Temporal Data with Tsibble and Fable"
author: "Justus v. Samson-Himmelstjerna & Oskar Krafft"
institute: "Hertie School"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    df_print: paged
    number_sections: FALSE
    highlight: tango
    theme: lumen
    toc_depth: 3
    toc_float: true
    css: magic/custom.css 
    self_contained: false
    includes:
      after_body: ../HTML_sources/footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

------------------------------------------------------------------------

### Workshop Packages

```{r, include = T}
library(tsibble)
library(fable)
library(tsibbledata)
```

### Supplementary Packages

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
```

<br>

------------------------------------------------------------------------

### Task 1 - Introduction: Create Tsibble objects

[2 points in total]

TASK DESCRIPTION

a)  Create a tsibble object that has "Year" as index and "Observation" key [0.5 points]

```{r}
my_year_vector <- 2010:2022
my_observation_vector <- rpois(n = 13, lambda = 50)

# Put your code here
y <- tsibble(
  Year = my_year_vector,
  Observation = my_observation_vector,
  index = Year,
  key = Observation
)

y
```

b)  coerce the `z` into a `tsibble` object, setting `Month` as index [1.5 points]
```{r}
z <- data.frame(
  Month = c("2022 Jan", "2022 Feb", "2022 Mar", "2022 Apr", "2022 May"),
  Observations = c(50, 23, 34, 30, 25)) %>% 
  mutate(Month = yearmonth(Month)) # correcting for the irregular amount of days per month

# Solution
z %>%
  as_tsibble(index = Month)
```

------------------------------------------------------------------------

### Task 2 - Forecasting Workflow: Data Preperation and Manipulation  

[10 points in total]

For the rest of the workshop, we will be using the `global_economy` dataset from the `tsibbledata` package.
The dataset includes economic indicators featured by the World Bank from 1960 to 2017.

```{r}
global_economy
```

a)  Identify which countries contain implicit gaps over the full time range [2 points]]

```{r}
# Solution
has_gaps(global_economy, .full = TRUE) %>% 
  filter(.gaps == TRUE)
```

b) compare your output to the output below and explain the different results [2 points]

```{r}
has_gaps(global_economy) %>% 
  filter(.gaps == TRUE)
```
**Solution**: here, has_gaps() only looks into the time range per key and not the entire time range


c) report the total number of implicit gaps in the dataset over the full time range [2 points]

```{r}
# Solution
global_economy %>% 
  count_gaps(.full = TRUE) %>% 
  summarise(gaps_total = sum(.n))
```

d) Fill all the identified gaps with `NA`s and show in an output that there are no more implicit gaps [2 points]

```{r}
#solution
tsbl_global_economy <- fill_gaps(global_economy, .full = TRUE)

tsbl_global_economy %>% 
  has_gaps(.full = TRUE) %>% 
  filter(.gaps == TRUE)
```

### Task 3 - Forecasting Workflow: Model specification and forecasting  

[18 points in total]

First, let's prepare the data by creating our variable of interest: `GDP_per_capita`

```{r}
gdp_pc <- tsbl_global_economy %>% #INSERT YOUR OBJECT FROM 1d between pipe operator and assignment operator. If you did not complete 1d) use the original global_economy tsibble
  mutate(GDP_per_capita = GDP / Population)
```

a) Using the `autoplot()` function, show the `GDP_per_capita` over time for China and Nigeria [6 points]

```{r}

# Solution
gdp_pc %>%
  filter(Country == c("Nigeria", "China")) %>%
  autoplot(GDP_per_capita) +
  labs(y = "$US", title = "GDP per capita for Nigeria and China")
```

b) Filter for China and Nigeria and only use the `Years` up to 2009 to train the models. Fit a default `arima` model and a `TSLM` model with a linear time `trend()` to `GDP_per_capita`. Finally, use `glance()` to inspect the models [6 points]

```{r}

# Solution
fit <- gdp_pc %>%
  filter(Country == c("Nigeria","China"),
         Year < 2010) %>% 
  model(arima = ARIMA(GDP_per_capita),
        lm = TSLM(GDP_per_capita ~ trend())
        )
glance(fit)
```


c) Now, forecast 7 years and `autoplot()` both forecasts. Provide an interpretation of the output. Which model creates a better forecast. Use time series vocabulary in your interpretation of the model performance [6 points]

```{r}

# Solution
fit %>%
  forecast(h = "7 years") %>%
  autoplot(gdp_pc, level = 80, alpha = 0.5) +
  labs(y = "$US", title = "GDP per capita for Nigeria and China")
```

